Simulating an Isolated Dwarf Galaxy with Energetic Feedback and Chemical Yields from Individual Stars

### 2018-03-15 main.tex
updated ArXiv references to include journal and status where available.

\subsubsection{Stellar Winds}
\label{sec:stellar winds}
Stellar winds are important sources of enrichment and feedback in galaxies at both early times from massive stars and late times from AGB stars. Although the energy injected via winds over the lifetime of a cluster of stars is much less than that from SNe and radiation \citep{Shull1995}, stellar winds are potentially important sources of pre-SN feedback. We inject the feedback energy in a spherical region with a three-cell radius centered on the emitting star particle, assuming full thermalization of the wind kinetic energy. The total injected energy in timestep $\Delta t$ is therefore $E_w = \dot{M}\Delta t \times v^2_w + E_{\rm th}$, where $E_{\rm th}$ is the thermal energy of the ejected gas mass $M_w = \dot{M}\Delta t$ given the star's interpolated effective temperature $T_{\rm eff}$. We use a Monte Carlo sampling method to inject mass and energy into the surrounding cells to account for partial cell overlap at the edge of the injection region that is a result of mapping a spherical injection region to cubic grid cells. We assume constant mass loss rates for all winds as set by the yield tables over either the lifetime of the star (for massive stars) or the length of the AGB phase (for low mass stars). 
%Properly modeling stellar winds requires sub-parsec resolution, which is problematic in galaxy-scale simulations.

@@

Stars that only undergo an AGB phase deposit their feedback at the end of their lives, as determined by the PARSEC evolution tracks. AGB wind velocities vary dramatically over their relatively short lifetimes, but are typically on the order of 10 km s$^{-1}$. For simplicity, we adopt a fixed wind velocity of 20 km s$^{-1}$ for all AGB stars.

\subsubsection{Core Collapse SNe}
\label{sec: core collapse}
Stars between  $M_{\rm thresh} = 8$ M$_{\odot}$ and 25 M$_{\odot}$ end their lives as core collapse SNe, ejecting mass and metals as determined by the NuGrid stellar yield tables, along with 10$^{51}$ erg of thermal energy. Due to the high resolution of our simulations (1.8~pc), we generally resolve the Sedov phase of each SN explosion well (see Appendix~\ref{appendix:SN}). We inject thermal energy alone in a three-cell radius spherical region around the star particle, which we find to be sufficient to resolve the SN explosions. We continue to track any remaining stellar mass after the SNe as a massive remnant tracer particle. In future work these particles can be used to self-consistently account for more exotic sources of feedback and chemical enrichment such as X-ray binaries and neutron-star binary merger events which, while rare, could be important in long term galaxy evolution \citep[e.g.][]{Artale2015}.

\subsubsection{Type Ia SNe}
\label{sec:Type Ia}
We continue to track low mass stars ($M < 8$ M$_{\odot}$) after their death as WD particles, marking a subset as Type Ia SN progenitors. This is the most self-consistent model for Type Ia SNe in galaxy scale simulations. We note however that for the low SFRs in our isolated dwarf galaxy simulation, the first Type Ia SN only appears after a few hundred megayears of simulation time. By the end of the simulation presented here (500~Myr), only 18 have gone off. At the end of their life, we assign a new mass to these particles following the initial-to-final-mass relation of \citet{Salaris2009}.
%mm \textbf{I originally reproduced the equations here, but it is currently commented out. - AE}
%or, 
%\[
%M_{\rm WD} = 
%\begin{cases}
%0.134 M_{\rm i} + 0.331 & 1.7~{\rm M_{\odot}} \le M_{\rm *} \le 4~{\rm M_{\odot}} \\
%0.047 M_{\rm i} + 0.679 & 4~{\rm M_{\odot}} \le M_{\rm *}
%\end{cases}
%\]
%where $M_{\rm *}$ is the initial mass of the WD progenitor.
We follow the common assumption that progenitor stars with initial masses between 3 and 8 M$_{\odot}$ form WDs that are Type Ia progenitors \citep[see][ and references therein]{Cote2017}.

@@

We also include the effects of radiation pressure on \ion{H}{1}. This has been shown to be important in suppressing the star formation rates of dwarf galaxies by influencing turbulence and the dense gas content of the ISM \citep{WiseAbel2012,Ceverino2014}. We ignore the absorption of ionizing radiation by dust and re-radiation in the infrared. This is included in other models \citep[e.g.][]{Rosdahl2015,FIRE,FIRE2} as this may increase by a factor of a few to several the effective radiation pressure \citep{ZhangDavis2017}. However, the importance of multiple scattering is still unclear. Other works have shown the effect to only increase the radiation pressure by a factor of order unity \citep{Krumholz2012,Krumholz2013,Reissl2017}. Due to these uncertainties, and given that our dwarf galaxy has a low dust content, and therefore a low infrared opacity, we ignore this effect.

>>>>

\subsubsection{Stellar Winds}
\label{sec:stellar winds}
Stellar winds are important sources of enrichment and feedback in galaxies at both early times from massive stars and late times from AGB stars. Although the energy injected via winds over the lifetime of a cluster of stars is much less than that from SNe and radiation \citep{Shull1995}, stellar winds are potentially important sources of pre-SN feedback. We inject the feedback energy in a spherical region with a three-cell radius centered on the emitting star particle, assuming full thermalization of the wind kinetic energy. The total injected energy in timestep $\Delta t$ is therefore $E_w = \dot{M}  v^2_w\Delta t  + E_{\rm th}$, where $E_{\rm th}$ is the thermal energy of the ejected gas mass $M_w = \dot{M}\Delta t$ given the star's interpolated effective temperature $T_{\rm eff}$. We use a Monte Carlo sampling method to inject mass and energy into the surrounding cells to account for partial cell overlap at the edge of the injection region that is a result of mapping a spherical injection region to cubic grid cells. We assume constant mass loss rates for all winds as set by the yield tables over either the lifetime of the star (for massive stars) or the length of the AGB phase (for low mass stars). 
%Properly modeling stellar winds requires sub-parsec resolution, which is problematic in galaxy-scale simulations.

@@

Stars that only undergo an AGB phase deposit their feedback at the end of their lives, as determined by the PARSEC evolution tracks. AGB wind velocities vary dramatically over their relatively short lifetimes, but are typically on the order of 10 km s$^{-1}$. For simplicity, we adopt a fixed wind velocity of 20 km s$^{-1}$ for all AGB stars as well.

\subsubsection{Core Collapse SNe}
\label{sec: core collapse}
Stars between  $
%mm3 M_{\rm thresh}
           M_{\rm trans} = 8$ M$_{\odot}$ and 25 M$_{\odot}$ end their lives as core collapse SNe, ejecting mass and metals as determined by the NuGrid stellar yield tables, along with 10$^{51}$ erg of thermal energy. Due to the high resolution of our simulations (1.8~pc), we generally resolve the Sedov phase of each SN explosion well (see Appendix~\ref{appendix:SN}). We inject thermal energy alone in a three-cell radius spherical region around the star particle, which we find to be sufficient to resolve the SN explosions.
%mm3 [How do you antialias the energy?  Something like Simpson's cloud-in-cell method, or something else?]
We continue to track any remaining stellar mass after the SNe occurs as a massive remnant tracer particle. In future work these particles can be used to self-consistently account for more exotic sources of feedback and chemical enrichment such as X-ray binaries and neutron-star binary merger events which, while rare, could be important in long term galaxy evolution \citep[e.g.][]{Artale2015}.

\subsubsection{Type Ia SNe}
\label{sec:Type Ia}
We continue to track low mass stars ($M < 8$ M$_{\odot}$) after their death as WD particles, marking a subset as Type Ia SN progenitors. This is the most self-consistent model for Type Ia SNe in galaxy-scale simulations. We note however that for the low SFRs in our isolated dwarf galaxy simulation, the first Type Ia SN only appears after a few hundred megayears of simulation time. By the end of the simulation presented here (500~Myr), only 18 have gone off. At the end of their life, we assign a new mass to these particles following the initial-to-final-mass relation of \citet{Salaris2009}.
%mm \textbf{I originally reproduced the equations here, but it is currently commented out. - AE}
%or, 
%\[
%M_{\rm WD} = 
%\begin{cases}
%0.134 M_{\rm i} + 0.331 & 1.7~{\rm M_{\odot}} \le M_{\rm *} \le 4~{\rm M_{\odot}} \\
%0.047 M_{\rm i} + 0.679 & 4~{\rm M_{\odot}} \le M_{\rm *}
%\end{cases}
%\]
%where $M_{\rm *}$ is the initial mass of the WD progenitor.
We follow the common assumption that progenitor stars with initial masses between 3 and 8 M$_{\odot}$ form WDs that are Type Ia progenitors \citep[see][ and references therein]{Cote2017}.

@@

We also include the effects of radiation pressure on \ion{H}{1}. This has been shown to be important in suppressing the star formation rates of dwarf galaxies by influencing turbulence and the dense gas content of the ISM \citep{WiseAbel2012,Ceverino2014}. We ignore the absorption of ionizing radiation by dust and re-radiation in the infrared. This is included in other models \citep[e.g.][]{Rosdahl2015,FIRE,FIRE2} as this may increase by a factor of a few to several the effective radiation pressure \citep{ZhangDavis2017}. However, the importance of multiple scattering is still unclear. Other works have shown the effect to only increase the radiation pressure by a factor of order unity \citep{Krumholz2012,Krumholz2013,Reissl2017,Wibking2018}. Due to these uncertainties, and given that our dwarf galaxy has a low dust content, and therefore a low infrared opacity, we ignore this effect.
